### 2023-02-15 chapters/conclusion.tex
conclusion modified

\chapter{Conclusion And Future Work} In the past decade, the advancement of machine learning has witnessed a host of  applications in different safety-critical domains. In such domains, the interpretability, fairness, robustness, and privacy etc.\ of classifiers are considered important in trustworthy and responsible AI (artificial intelligence). In this thesis, we focus on the interpretability and fairness aspects of machine learning, where we prioritize on improving the scalability and accuracy of the underlying problems. Relying on formal methods, our contributions are summarized as follows: (i) in interpretable machine learning, we design incremental learning technique for interpretable rule-based classifiers, (ii) in fairness in machine learning, we develop a formal fairness verification framework to verify multiple fairness definitions, algorithms, and classifiers; and we develop techniques to identify the sources of unfairness of classifiers by computing fairness influence functions.

To validate the efficiency of our methods, we develop open-source tools and conduct experiments on real-world interpretability and fairness datasets in machine learning. In interpretable machine learning, our framework {\imli} scales classification to million-size datasets while achieving competitive prediction accuracy and rule-size compared to existing rule-based classifiers. In our quest for learning more expressible interpretable classifiers, we develop another learning framework {\crr} based on incremental learning. In experiments, {\crr} learns more concise and accurate rule-based classifiers while demonstrating scalability owing to incremental learning.

In fairness in machine learning, our probabilistic fairness verifier {\justicia} and {\fvgm} demonstrates a superior performance in accuracy and scalability than the state of the art verifiers. Being a stochastic-SAT-based verifier, {\justicia} verifies fairness of compound sensitive groups of Boolean classifiers such as decision trees with higher scalability. {\fvgm}, in addition, considers correlated features and verifies fairness of linear classifiers with better scalability and accuracy than existing verifiers. Finally, given a classifier and a dataset, our framework {\fairXplainer} computes fairness influences of each subset of non-sensitive features as their contribution to the incurred bias of the classifier. In experiments, {\fairXplainer} estimates bias with significantly higher accuracy than the existing local explanation approach, while enabling computation of intersectional influences as a mean of better understanding the sources fo bias.

>>>>

\chapter{Conclusion And Future Work}
\label{chapter:conclusion} 
Over the past decade, machine learning has been applied to various safety-critical domains, and it's crucial for classifiers to be interpretable, fair, robust, and private to ensure trustworthy and responsible AI. In this thesis, we focus on the interpretability and fairness aspects of machine learning and aim to improve the scalability and accuracy of the underlying problems. We utilize formal methods to make the following contributions: (i) In interpretable machine learning, we design an incremental learning technique for interpretable rule-based classifiers of varied expressiveness. (ii) In fairness in machine learning, we develop a formal probabilistic fairness verification framework that can verify multiple fairness definitions of classifiers.  Additionally, we develop techniques to explain fairness metrics by identifying feature combinations responsible for the bias of the classifier.

To demonstrate the efficacy of our methods, we have constructed open-source tools and conducted experiments on real-world datasets in machine learning to evaluate interpretability and fairness. In the context of interpretable rule-based machine learning, we have developed an incremental learning framework, known as {\imli}, that scales classification to million-size datasets while maintaining competitive prediction accuracy and rule-size compared to existing rule-based classifiers. Additionally, in our pursuit of more expressive yet interpretable classifiers, we have introduced another learning framework, called {\crr}, for logical relaxed classification rules that are based on incremental learning. Experimental results show that {\crr} is capable of learning more concise and accurate rule-based classifiers.

Our work on fairness in machine learning introduces two novel probabilistic fairness verifiers, {\justicia} and {\fvgm}, which exhibit superior performance in accuracy and scalability compared to state-of-the-art verifiers. {\justicia} is a stochastic-SAT-based verifier that enables scalable verification of fairness for compound sensitive groups of Boolean classifiers, such as decision trees, which was previously infeasible with existing methods. On the other hand, {\fvgm} takes correlated features into account and is capable of verifying the fairness of linear classifiers with higher scalability and accuracy than previous verifiers. Additionally, we propose a global sensitivity analysis-based method, {\fairXplainer}, that explains group fairness metrics by computing fairness influences of individual and intersectional features. Notably, {\fairXplainer} approximates bias more accurately using fairness influence functions (FIFs) and demonstrates a higher correlation of FIFs with fairness intervention than the local explainability-based approach.

In future, our research on fair and interpretable machine learning will focus on several areas. One immediate research question is to design a fairness improvement algorithm for machine learning classifiers using the fairness verifiers, {\justicia} and {\fvgm}, and the bias explanation framework, {\fairXplainer}. We also aim to extend formal fairness verification to more complex classifiers, such as random forests and neural networks. In addition, our future plans include deploying {\fairXplainer} on image classifiers, language models, and computer vision tasks. In a separate line of work, we will explore the application of incremental solving, as demonstrated in the context of interpretable classification learning by {\imli} and {\crr}, to other optimization problems, even beyond machine learning. To conclude the long-term vision of this thesis is to enhance machine learning in interpretability, fairness, and beyond, with the support of formal methods and automated reasoning.

\begin{comment}
	\begin{itemize}
		\item Fairness repair
		\item Incremental solving
		\item Fairness and interpretability as a service to more complex models.
	\end{itemize}
\end{comment}
